---
title: "Cherry_Picker"
author: "Gianluca Pegoraro"
date: "December, 15 2016"
output: html_document
---
## Aim

This script uses as input a .csv file containing the number and positions of "Positive" wells to be picked from an arbitrary number of 96-well source plates. This table is then rearranged to produce a .csv file containing the well positions for a matched number of 96-well destination plates. The script allows for 2 output formats: `default`- which included the whole 96-wells on the destination plates -, or `internal` - which excludes all the edge wells on the plates. This behavior can be controlled by the `format` flag to be specified at the beginning of the script run. 

```{r loadAndClean}
library(plyr)
library(ggplot2)
library(data.table)
library(stringr)
library(knitr)
```

```{r, include=FALSE, warning=FALSE}
opts_chunk$set(
    fig.path = "output/",
    cache = TRUE,
    dev = c("png"),
    message = FALSE,
    warning = FALSE
    )
    theme_set(theme_minimal())
```

Set the import flag. The Flag variable can only take the `default` or `internal` values. Also, set the desired number of source plates for the template file.
```{r inputs}
flag <- "internal"

n_source <- 20
```

## Generate a template input file

Generate some template data (Full source plate).
```{r}
set.seed(161115)

input_template <-
    data.table(
    source_rack = rep(paste0("Input_Rack_", 1:n_source), each = 96),
    source_id =  paste0("ID_", sample(1:10000, 96 * n_source)),
    source_pos = rep(1:96, n_source),
    source_col = rep(1:12, each = 8),
    source_row = rep(1:8, 12),
    source_flag = sample(
    c("Hit", "Miss"),
    96 * n_source,
    replace = TRUE,
    prob = c(0.35, 0.65)
    )
    )

input_template[, `:=`(source_well = paste0(LETTERS[source_row], source_col))]
```

Randomize empty wells in the Source plates (`Miss`)
```{r}
input_template[source_flag == "Miss", source_id := NA]
```

Write the template table to a .csv file.
```{r}
write.csv(input_template,
          "Input_template.csv",
          row.names = FALSE,
          quote = FALSE)
```

The `Input_template.csv` can be hand modified in excel. Edit the content of single cells as needed, but **Do not modify the number or name of the variables (Columns) in the file**. These are the variables (Columns) in the file that should be present:

* `source_rack`: The barcode (name) of the source plate | Required 
* `source_id`: An identifier for a the source well | Optional
* `source_pos`: A numerical identifier for the well position | Required (**Do not edit**)
* `source_col`: A numerical identifier for the well column | Required (**Do not edit**)
* `source_row`: A numerical identifier for the well row | Required (**Do not edit**)
* `source_well`: A string name for the well position | Required (**Do not edit**)
* `source_flag`: A flag indicating whether the well should be transferred or not| Required (**Can assume only the `Hit` or `Miss` values)

Once the contents of the files are edited according to your needs, save the file in .csv format with the same name.

## Read and process the modified input file
Read the .csv input file. 
```{r}
input <- fread("Input_template.csv")
```

Plot the Source plates layout
```{r, fig.height=12, fig.width= 18}
input_plot <- ggplot(input, aes(x = source_col, y = source_row, label = source_id))
input_plot + geom_text(size = 1.5) +
            facet_wrap( ~ source_rack) +
            scale_x_continuous(lim = c(1, 12), breaks = 1:12) +
            scale_y_reverse(lim = c(8, 1), breaks = 1:8, labels = LETTERS[1:8]) +
            ggtitle("Source Plates Layout")
```

Filter out empty wells to generate the output table. Also count how many source plates are actually present. 
```{r}
output <- input[source_flag == "Hit",]

n_source <- length(output[,unique(source_rack)])
```

Assign the destination positions. Only the internal 60 wells (Excluding rows A/H and columns 1/12) are used. 
```{r}
try(if(!(flag %in% c("default", "internal"))) stop("Invalid format flag, set it either to 'default' or 'internal'", call. = FALSE))

if(flag == "default"){
    output[, `:=`(dest_rack = rep(paste0("Output_Rack_", 1:n_source), each = 96),
                 dest_pos = 1:96,
                 dest_col = rep(1:12, each = 8),
                 dest_row = rep(1:8, 12))]
    
    output[, `:=`(dest_well = paste0(LETTERS[dest_row], dest_col))]
} 

if(flag == "internal") {
    output[, `:=`(dest_rack = rep(paste0("Output_Rack_", 1:(n_source*10)), each = 60),
                 dest_pos = as.vector(outer((10:15), seq(0, 72, 8), FUN = "+")),
                 dest_col = rep(2:11, each = 6),
                 dest_row = rep(2:7, 10))]
    
    output[, `:=`(dest_well = paste0(LETTERS[dest_row], dest_col))]
}
```

Plot the Source plates layout
```{r}
output_plot <- ggplot(output, aes(x = dest_col, y = dest_row, label = source_id))
output_plot + geom_text(size = 1.5) +
            facet_wrap( ~ dest_rack) +
            scale_x_continuous(lim = c(1, 12), breaks = 1:12) +
            scale_y_reverse(lim = c(8, 1), breaks = 1:8, labels = LETTERS[1:8]) +
            ggtitle("Dest Plates Layout")
```

```{r}
write.csv(x = output, 
          file = "output/Cherry_Pick_Worklist.csv", 
          quote = FALSE, 
          col.names = TRUE,
          row.names = FALSE)
```
Document the information about the analysis session
```{r sessionInfo, include=TRUE, echo=TRUE, results='markup'}
sessionInfo()
```